---
title: "R Notebook"
output: html_notebook
---

## retrieve and prepare the previously cleaned csv file (Merged_County_Sales.csv)

```{r}
library(stringr)
library(dplyr)
library(formattable)
library(ggplot2)

#retrieve manually cleaned file
merged_county_sales <- read.csv("C:/Users/David/Desktop/MSDS696/Colorado MJ Datasets/Merged_County_Sales.csv", header=TRUE, colClasses = c("character", "character", "factor", "character", "character"))

#convert date field to R recognized date format
merged_county_sales$Date <- as.Date(merged_county_sales$Date)

#convert currency fields from factor to numeric
merged_county_sales$Med_County_Sales <- as.numeric(as.character(merged_county_sales$Med_County_Sales))
merged_county_sales$Rec_County_Sales <- as.numeric(as.character(merged_county_sales$Rec_County_Sales))

#convert currency fields from numeric to currency
merged_county_sales$Med_County_Sales <- currency(merged_county_sales$Med_County_Sales, symbol="$", big.mark=",", digits = 0L)
merged_county_sales$Rec_County_Sales <- currency(merged_county_sales$Rec_County_Sales, symbol="$", big.mark=",", digits = 0L)

summary(merged_county_sales)

summary(merged_county_sales$Rec_County_Sales)

options(scipen=10)

ggplot(data=merged_county_sales, mapping=aes(x=Date))+
  geom_bar(mapping=aes(y=Rec_County_Sales/100000), stat="identity", position="dodge", fill="green")+
  geom_bar(mapping=aes(y=Med_County_Sales/100000), stat="identity", position="dodge", fill="red")+
  labs(title="Colorado Medical and Retail Marijuana Sales 2014-2018", x= "Years of Recreational Legalization", y= "Sales per $100K")+
  theme_bw()

ggplot(data=merged_county_sales, aes(x=as.factor(merged_county_sales$County)))+
  geom_boxplot(aes(y=Rec_County_Sales), color="green")+
  geom_boxplot(aes(y=Med_County_Sales), color="red")+
  labs(title="Colorado Retail and Medical Marijuana Sales since 2014", x= "Colorado County", y= "Sales")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

### investigate county sales by year
Year <- as.integer(paste(word(merged_county_sales$Date, 1, sep="-")))
merged_county_sales <- cbind(merged_county_sales, Year)

#x <- group_by(merged_county_sales, Year)
#x <- summarize(Year, Med_County_Sales = sum(Med_County_Sales))
#x


#write.csv(merged_county_sales, file = "merged_county_sales.csv")

```


```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(ggrepel)

colopop_df <- as.data.frame(read.csv("C:/Users/David/Desktop/MSDS696/Colorado Demographics/CountyPopulation.csv", col.names=c("CountyFIPS", "Area", "Year", "Population"), colClasses = c("character", "factor", "factor", "numeric")))

state_pop <- filter(colopop_df, Area == "COLORADO STATE")
county_pop <- filter(colopop_df, Area != "COLORADO STATE")

summary(state_pop)

options(scipen=10)

#View(state_pop)

ggplot(data=state_pop, mapping=aes(x=Year, y=Population, group=Area))+
  geom_line(color="red")+
  geom_point(color="red")+
  scale_y_continuous(limits = c(min(state_pop$Population), max(state_pop$Population)))+
   geom_label_repel(aes(label = Population), nudge_x = 1)+
  labs(title="Colorado Estimated Population Growth since 2010 Census")+
  theme_bw()
 
```

```{r}
#retrieve manually cleaned file
#library(readr)
#library(dplyr)
#library(ggplot2)

#co_pop <- as.data.frame(read.csv("C:/Users/David/Desktop/MSDS696/Colorado Demographics/CountyRankings.csv", col.names=c("CountyFIPS", "County", "pop_2010", "pop_2011", "pop_2012", "pop_2013", "pop_2014", "pop_2015", "pop_2016", "pop_2017", "rank_2017"), colClasses = c("character", "factor", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "integer")))

#state_pop <- co_pop[1, -1]
#county_pop <- co_pop[-1, -1]

#reorder by city population rank (2017)
#county_pop <- county_pop[order(county_pop$rank_2017), ]
#county_pop <- county_pop[1:20, ]

#county_pop.f1 <- filter(county_pop, pop_2010 >= 100000)
#county_pop.f2 <- filter(county_pop, pop_2010 < 100000 & pop_2010 > 25000)
#county_pop.f3 <- filter(county_pop, pop_2010 < 25000)

#rm(co_pop)

#View(county_pop)

#transpose_df <- as.data.frame(t(county_pop))
#View(county_pop)

#options(scipen=10)

#ggplot(data=county_pop.f1, mapping=aes(x=County))+
  #geom_col(mapping=aes(y=pop_2017, fill="2017"), position = "dodge")+
  #geom_col(mapping=aes(y=pop_2014, fill="2014"), position = "dodge")+
  #geom_col(mapping=aes(y=pop_2010, fill="2010"), position = "dodge")+
  #labs(title="", x= "Colorado County", y= "Sales per $1M")+
  #theme_bw()+
  #theme(axis.text.x = element_text(angle = 90, hjust = 1))

#ggplot(data=county_pop.f2, mapping=aes(x=County))+
  #geom_col(mapping=aes(y=pop_2017, fill="2017"), position = "dodge")+
  #geom_col(mapping=aes(y=pop_2014, fill="2014"), position = "dodge")+
  #geom_col(mapping=aes(y=pop_2010, fill="2010"), position = "dodge")+
  #labs(title="", x= "Colorado County", y= "Sales per $1M")+
  #theme_bw()+
  #theme(axis.text.x = element_text(angle = 90, hjust = 1))

#ggplot(data=county_pop.f3, mapping=aes(x=County))+
  #geom_col(mapping=aes(y=pop_2017, fill="2017"), position = "dodge")+
  #geom_col(mapping=aes(y=pop_2014, fill="2014"), position = "dodge")+
  #geom_col(mapping=aes(y=pop_2010, fill="2010"), position = "dodge")+
  #labs(title="", x= "Colorado County", y= "Sales per $1M")+
  #theme_bw()+
  #theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
library(ggrepel)
library(ggplot2)

co_sin_tax <- as.data.frame(read.csv("C:/Users/David/Desktop/MSDS696/colorado_sin_revenue.csv"))

summary(co_sin_tax)

options(scipen=10)

#create a dataframe with the sin tax and co population data. the data will need to be normalized.

df <- co_sin_tax
df <- merge(co_sin_tax, state_pop)

df$med_marijuana_tax <- currency(df$med_marijuana_tax, symbol="$", big.mark=",", digits = 0L)
df$rec_marijuana_tax <- currency(df$rec_marijuana_tax, symbol="$", big.mark=",", digits = 0L)
df$alcohol_tax <- currency(df$alcohol_tax, symbol="$", big.mark=",", digits = 0L)
df$tobacco_tax <- currency(df$tobacco_tax, symbol="$", big.mark=",", digits = 0L)
df$gambling <- currency(df$gambling, symbol="$", big.mark=",", digits = 0L)

ggplot(data=df, mapping=aes(x = Year)) +
  geom_point(aes(y=rec_marijuana_tax/100000, color="Recreational Marijuana"))+
  geom_line(aes(y=rec_marijuana_tax/100000, color="Recreational Marijuana"), size=2)+
  geom_point(aes(y=alcohol_tax/100000, color="Alcohol"))+
  geom_line(aes(y=alcohol_tax/100000, color="Alcohol"))+
  geom_point(aes(y=tobacco_tax/100000, color="Tobacco"))+
  geom_line(aes(y=tobacco_tax/100000, color="Tobacco"))+
  geom_point(aes(y=gambling/100000, color="Gambling"))+
  geom_line(aes(y=gambling/100000, color="Gambling"))+
  geom_label_repel(aes(y=rec_marijuana_tax/100000, label=rec_marijuana_tax), nudge_x = 0, nudge_y = -200)+
  scale_colour_manual("", 
                      breaks = c("Medical Marijuana", "Recreational Marijuana", "Alcohol", "Tobacco", "Gambling"),
                      values = c("Medical Marijuana"="green", "Recreational Marijuana"="red", 
                                 "Alcohol"="purple", "Tobacco"="blue", "Gambling"="orange"))+
  labs(title="Comparisons of Colorado 'sin' Taxes 2013-2017", x= "Year", y= "Colorado Marijuana Revenue per $100K")+
  theme_bw()+
  theme(legend.position="top")

###perform correlation analysis - I don't think really tells me anything of value. 

#scale the data
#df_year <- df[1]
#df_scaled <- df[c(2:6,9)]
#df_scaled <- cbind(df_year, df_scaled)

#library(corrplot)
#m <- cor(df_scaled[-1, 2:6])
#corrplot(m, type = "lower", order = "hclust", method="number")

View(df)


```

```{r}

library(stringr)
library(dplyr)
library(formattable)
library(ggplot2)

merged_county_sales2 <- read.csv("C:/Users/David/Desktop/MSDS696/Colorado MJ Datasets/Merged_County_Sales_v2.csv", header=TRUE)
#merged_county_sales2 <- as.data.frame(read.csv("C:/Users/David/Desktop/MSDS696/Colorado MJ Datasets/Merged_County_Sales_v2.csv", header=TRUE, colClasses = c("character", "character", "factor", "numeric", "numeric", "integer", "numeric", "numeric", "numeric")))

merged_county_sales2$Med_County_Sales <- currency(merged_county_sales2$Med_County_Sales, symbol="$", big.mark=",", digits = 0L)
merged_county_sales2$Rec_County_Sales <- currency(merged_county_sales2$Rec_County_Sales, symbol="$", big.mark=",", digits = 0L)

summary(merged_county_sales2)

ggplot(data=merged_county_sales2, mapping=aes(x=Year))+
  geom_col(mapping=aes(y=Rec_County_Sales/100000), fill="green")+
  geom_col(mapping=aes(y=Med_County_Sales/100000), fill="red", alpha=.9)+
  labs(title="Colorado Medical and Retail Marijuana Sales 2014-2018", x= "Years of Recreational Legalization", y= "Sales per $100K")+
  theme_bw()

#test <- merged_county_sales2[order(merged_county_sales2$Rec_County_Sales, merged_county_sales2$County),] 

ggplot(data=merged_county_sales2, aes(x=County))+
  geom_boxplot(aes(y=Rec_County_Sales/100000), color="green")+
  geom_boxplot(aes(y=Med_County_Sales/100000), color="red")+
  labs(title="Colorado Retail and Medical Marijuana Sales since 2014", x= "Colorado County", y= "Sales")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90))

ggplot(data=merged_county_sales2, mapping=aes(x = Year)) +
  geom_col(aes(y=County_Population, color="County_Population"))

View(merged_county_sales2)



```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
